---
title: "Exam II R"
author: "Filip Wilhelm Sjostrand"
date: "18/11-22"
output: pdf_document
---

```{r message=FALSE, echo=FALSE}
# Packages ----------
library(readr)
library(ggplot2)
library(dplyr)

# Get data ----------
df <- read_csv(
  file = "~/Downloads/athelete.csv",
  col_types = cols(
    sex = col_factor(levels = c("m", "f")),
    newsport = col_factor(levels = c("Net", "Swim", "Run"))
    )
  )



```

\newpage

# Problem 1

### (a) Write down the estimated linear regression model for the full model.

```{r, echo=FALSE}
# -------------------- QUESTION I --------------------

# Full model ----------
model <- lm(formula = rcc ~ lbm + bmi + pcBfat + ferr + sex + newsport, data=df)

b0 <- signif(model$coefficients[1], 4)
b1 <- signif(model$coefficients[2], 4)
b2 <- signif(model$coefficients[3], 4)
b3 <- signif(model$coefficients[4], 4)
b4 <- signif(model$coefficients[5], 4)
b5 <- signif(model$coefficients[6], 4)
b6 <- signif(model$coefficients[7], 4)
b7 <- signif(model$coefficients[8], 4)

```

$$\hat{Y}= `r b0` + `r b1`X_1 + `r b2`X_2 `r b3`X_3 `r b4`X_4 `r b5`X_{5,f} + `r b6`X_{6,S} + `r b7`X_{6,R}$$
\


### (b) Interpret $b_1$ and $b_6$ in terms of the problem.

**$b_1$:** Holding all other variables constant, when lean body mass increase by 1 kg, the red blood cell count increase by `r b1` per liter, on average.

**$b_6$:** Holding all other variables constant, the average difference between a net sports player and swim sports player is `r b6` per liter of red blood cell count.

\

### (c) Test to see if we can drop the $X_1$, $X_2$, and $X_3$ from the model. State the hypotheses in terms of the betas, test statistic, p-value, and conclusion.

Hypothesis: $H_0:$ $\beta_1 = \beta_2 = \beta_3 = 0$ VS. $H_a:$ at least one $\beta_i \ne 0$

```{r, echo=FALSE}
# Full vs reduced ----------
reduced <- lm(formula = rcc ~ ferr + sex + newsport, data=df)
sse_r <- anova(reduced)$`Sum Sq`[4]
df_r <- anova(reduced)$Df[4]

sse_f <- anova(model)$`Sum Sq`[7]
df_f <- anova(model)$Df[7]

Fstat <- ( (sse_r - sse_f) / (df_r - df_f) ) / (sse_f / df_f)

pvalue <- pf(Fstat, (df_r - df_f), df_f, lower.tail = FALSE)

```

Test statistic: $F = `r signif(Fstat, 4)`$

P-value$= `r signif(pvalue, 4)`$

Conclusion: at 5% significance level we fail to reject $H_0$ and conclude that there is insufficient evidence that $X_1$, $X_2$, and $X_3$ has significant slopes. Thus, they can be dropped from the model.

\

### (d) Based on your results from part (c), test to see if we can drop $X_6$ from the "best" model. State the hypotheses in terms of the betas, test statistic, p-value, and conclusion.

Hypothesis: $H_0:$ $\beta_6 = \beta_7 =  0$ VS. $H_a:$ at least one $\beta_i \ne 0$

```{r, echo=FALSE}
# Full vs reduced ----------
full <- lm(formula = rcc ~ ferr + sex + newsport, data=df)
sse_f <- anova(full)$`Sum Sq`[4]
df_f <- anova(full)$Df[4]

reduced <- lm(formula = rcc ~ ferr + sex, data=df)
sse_r <- anova(reduced)$`Sum Sq`[3]
df_r <- anova(reduced)$Df[3]

Fstat <- ( (sse_r - sse_f) / (df_r - df_f) ) / (sse_f / df_f)

pvalue <- pf(Fstat, (df_r - df_f), df_f, lower.tail = FALSE)

```

Test statistic: $F = `r signif(Fstat, 4)`$

P-value $= `r signif(pvalue, 4)`$

Conclusion: at 5% significance level we reject $H_0$ and conclude that there is sufficient evidence that $\beta_6$ and/or $\beta_7$ has a significant slope. Thus, they cannot be dropped from the model.

\

### (e) What is the additional reduction in error we expect to see when we add $X_5$ to a model which already includes $X_4$ and $X_6$. 

```{r, echo=FALSE}
# Partial r-squared ----------
before <- lm(formula = rcc ~ ferr + newsport, data=df)
sse_b <- anova(before)$`Sum Sq`[3]

after <- lm(formula = rcc ~ ferr + sex + newsport, data=df)
sse_a <- anova(after)$`Sum Sq`[4]

part_r <- (sse_b - sse_a)/sse_b



```

Adding sex to a model which already includes plasma ferritins and sport reduces error by `r round(part_r*100, 2)` %.
\newpage

# Problem 2

### (a) Based on your "best" model above, predict the red blood cell count of a male subject who a ferritin level of 20ng and swims.

```{r, echo=FALSE}
# -------------------- QUESTION II --------------------

# Prediction ----------
best <- lm(formula = rcc ~ ferr + sex + newsport, data=df)
newdata <- data.frame(ferr = 20, sex = "m", newsport = "Swim")
pred <- predict(best, newdata)

```

Answer: `r round(pred, 2)` per liter.

\

### (b) Find the 95\% confidence interval for only the value $\beta_1$, and interpret it in terms of the problem.

```{r, echo=FALSE}
# Confidence interval ----------
ci = confint(best, level = 0.95)

```

We are 95% confident that the true slope for Plasma ferritins is between `r round(ci[2,1], 4)` and `r format(round(ci[2,2], 4), scientific = FALSE)`. Thus, we cannot conclude that the slope is different from zero.

\

### (c) Create two prediction intervals for the red blood cell count for a male who runs with a ferratin level of 15 and a female who swims with a ferratin level of 12 with overall/simultaneous/family-wise level 90%.  


```{r,echo=FALSE}
# Predictions----------
newdata = data.frame(
  ferr = c(15, 12),
  sex = c("m", "f"), 
  newsport = c("Run", "Swim")
  )

preds <- predict(best, newdata)

# Find best multiplier ----------
Multiplier = function(n,p,g,alpha){
  Bon = qt(1-alpha/(2*g), n-p)
  WH = sqrt(p*qf(1-alpha,p,n-p))
  Sch = sqrt(g*qf(1-alpha,g, n-p))
  the.multipliers = round(c(Bon,WH,Sch), 4)
  names(the.multipliers) = c("Bonferroni", "WH", "Scheffe")
  return(the.multipliers)
}

bon <- Multiplier(nrow(df), 5, 2, 0.1)[1] #Bonferroni was the smallest 

# Error ----------
se = predict(best, newdata, interval = 'prediction', se.fit = TRUE)$se.fit
MSE = sum(best$residuals^2)/ (length(best$residuals) - length(best$coefficients))
pred.se = sqrt(se^2 + MSE)

# Prediction intervals ----------
intervals <- cbind(preds - bon*pred.se, preds + bon*pred.se)
colnames(intervals) <- c("2.5%", "97.5%" )
rownames(intervals) <- c("Male, runs, 15ng", "Female, swims, 12ng")



```
`r knitr::kable(intervals)`

\

### (d) Interpret the intervals in (c) in terms of the problem.

We are overall 90% confident that a male who runs with 15 in ferratin level has a blood plasma count between `r round(intervals[1,1],3)` and `r round(intervals[1,2],3)` per liter, and a woman who swims with 12 in ferratin level has a bloom plasma count between `r round(intervals[2,1],3)` and `r round(intervals[2,2],3)` per liter.

\newpage

# Problem 3

### (a) Fit the model which includes $X_3$ and $X_6$ and all interaction terms. Report back the estimated regression line. 

```{r, echo=FALSE}
# -------------------- QUESTION III --------------------

# Interaction model ----------
int_model <- lm(formula = rcc ~ pcBfat*newsport, data=df)
coef <- int_model$coefficients
b0 <- signif(coef[1], 4)
b1 <- signif(coef[2], 4)
b2 <- signif(coef[3], 4)
b3 <- signif(coef[4], 4)
b4 <- signif(coef[5], 4)
b5 <- signif(coef[6], 4)

```

$$\hat{Y} = `r b0` `r b1`X_3 + `r b2`X_{6,S} `r b3`X_{6,R} `r b4`X_{3}X_{6,S} + `r b5`X_{3}X_{6,R}$$

\

### (b) Test to see if you can drop the interaction term from the model in (a), using $\alpha=0.10$ State the null and alternative in terms of the betas, the p-value, and your conclusion in terms of the problem.

Hypothesis: $H_0:$ $\beta_4 = \beta_5 =  0$ VS. $H_a:$ at least one $\beta_i \ne 0$

```{r, echo=FALSE}
# Test of interaction ----------
noint_model <- lm(formula = rcc ~ pcBfat + newsport, data=df)
result <- anova(noint_model, int_model)
pval <- result$`Pr(>F)`[2]

```
p-value $=`r round(pval,4)`$

Conclusion: Given that p-value $< \alpha$, we reject $H_0$. Thus we conclude at 10% significance level that there is sufficient evidence that the interaction terms should not be dropped.

\

### (c) Using ggplot2, plot the different lines suggested by the interaction term on the plot with $Y$ and $X_3$.

```{r, echo=FALSE, message=FALSE}
int_model %>%
  ggplot(aes(x = pcBfat, y = rcc, group = newsport, color = newsport)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  ylab("Red blood cell count (per liter)") +
  xlab("Percent body fat") +
  labs(color = 'Sport') +
  labs(title = "Linear Regression with Interaction Terms")
 

```

\

### (d) State the different models associated with the different levels of $X_6$. Explain the relationship.

**Net:** $\hat{Y} = `r b0` `r b1`X_3$

**Swim:** $\hat{Y} = `r b0+b2` `r b1+b4`X_3$

**Run:** $\hat{Y} = `r b0 + b3` `r b1 + b5`X_3$

There is a negative relationship between body fat percentage and red blood cell count. On average, holding body fat constant, swimmers have a higher count of red blood cells compared to runners and net players, and runners have a higher than net players. However, per increase in body fat percentage, swimmers are expected to have the biggest decrease in red blood cell count on average while runners are expected to have the smallest.

\

### (e) Find and interpret $R^2\{X_3X_6| X_3,X_6\}$

```{r, echo=FALSE}
# Partial r-square ----------
sse_b <-  sum(summary(noint_model)$residuals^2)
sse_a <-  sum(summary(int_model)$residuals^2)
r <-  signif(((sse_b - sse_a)/sse_b)*100, 4)

```
When we add interaction terms to a model that already contain body fat percentage and sports, the error is reduced by `r r`%.

\newpage

# Code Appendix
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```